---
title: "H~4~: Funding sources"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(scales)
library(ggh4x)
library(glue)
library(gt)
library(gtExtras)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

preds_all <- tar_read(preds_conditional_treatment_only)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

# H~4a~: Funding sources

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 4~a~: Funding sources

**Donors will show increased willingness to donate to NGOs that are funded primarily by numerous small private donors compared to NGOs that are funded by a handful of wealthy private donors and government grants**

*Mechanisms: Perception of efficacy - your contribution matters as a small donor. Government funding may also imply lack of independence of government which can reduce the efficiency of an organization.*
:::

## Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Few wealthy donors or Government} \right) \bigr] - \\
&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Many small donors} \right) \bigr]
\end{aligned}
$$

```{r h4a-calc-estimands}
h4a_mms <- preds_all %>% 
  group_by(feat_funding_short, .draw) %>% 
  summarize(avg = mean(.epred))

h4a_amces <- h4a_mms %>% 
  group_by(feat_funding_short) %>% 
  compare_levels(variable = avg, by = feat_funding_short, comparison = "control")
```

```{r plot-h4a-mm-amce, fig.width=10, fig.height=4, out.width="100%"}
#| column: body-outset-right
p_h4a_mms <- h4a_mms %>% 
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  )

p_h4a_amces <- h4a_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_funding_short,
    delim = " - ", 
    names = c("feat_funding_short", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_funding_short = unique(.$reference_level)) %>%
  mutate(feat_funding_short = factor(feat_funding_short, levels = levels(h4a_mms$feat_funding_short))) %>% 
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Average marginal component effects (AMCEs)"
  )

p_h4a_mms | p_h4a_amces
```

```{r table-h4a-mm-amce}
#| column: body-outset-right
h4a_tbl_mm <- h4a_mms %>% 
  group_by(feat_funding_short) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_funding_short)) %>% 
  select(feat_funding_short, nice)

h4a_tbl_amces <- h4a_amces %>% 
  group_by(feat_funding_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_funding_short = str_replace(feat_funding_short, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_funding_short)) %>% 
  select(contrast = feat_funding_short, amce_nice = nice, p_neq_0)

bind_cols(
  h4a_tbl_mm, 
  add_row(h4a_tbl_amces, contrast = "*(Reference)*")
) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, amce_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_funding_short, contrast)) %>% 
  cols_label(
    feat_funding_short = "Funding source", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    amce_nice = "Posterior AMCE",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, amce_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```


# H~4b~: Relationship with host government and funding sources

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 4~b~: Relationship with host government and funding sources

**Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors**
:::

## Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Few wealthy donors or Government} \mid \text{Relationship = Under crackdown or Criticized} \right) \bigr] - \\
&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Many small donors} \mid \text{Relationship = Under crackdown or Criticized} \right) \bigr]
\end{aligned}
$$

```{r h4b-calc-estimands}
h4b_mms <- preds_all %>% 
  group_by(feat_funding_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

h4b_amces <- h4b_mms %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_funding_short, comparison = "control")
```

```{r plot-h4b-mm-amce, fig.width=10, fig.height=6.5, out.width="100%"}
#| column: body-outset-right
p_h4b_mms <- h4b_mms %>% 
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  facet_wrap(vars(feat_govt), ncol = 1) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )

p_h4b_amces <- h4b_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_funding_short,
    delim = " - ", 
    names = c("feat_funding_short", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_funding_short = "Many small donors", feat_govt = "Friendly relationship with government") %>%
  add_row(avg = 0, feat_funding_short = "Many small donors", feat_govt = "Criticized by government") %>%
  add_row(avg = 0, feat_funding_short = "Many small donors", feat_govt = "Under government crackdown") %>%
  mutate(feat_funding_short = factor(feat_funding_short, levels = levels(h4b_mms$feat_funding_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h4b_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  facet_wrap(vars(feat_govt), ncol = 1) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h4b_mms | p_h4b_amces
```

```{r table-h4b-mm-amce}
#| column: body-outset-right
h4b_tbl_mm <- h4b_mms %>% 
  group_by(feat_govt, feat_funding_short) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_govt), desc(feat_funding_short)) %>% 
  select(feat_funding_short, feat_govt, nice)

h4b_tbl_amces <- h4b_amces %>% 
  group_by(feat_govt, feat_funding_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  ungroup() %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_funding_short = str_replace(feat_funding_short, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_funding_short, diff_nice = nice, p_neq_0)

bind_cols(
  h4b_tbl_mm, 
  h4b_tbl_amces %>% 
    add_row(contrast = NA, .after = 2) %>% 
    add_row(contrast = NA, .after = 5) %>% 
    add_row(contrast = NA, .after = 8)
) %>% 
  mutate(feat_govt = fct_relabel(feat_govt, ~paste("Relationship with government:", .x))) %>% 
  group_by(feat_govt) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, diff_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_funding_short, contrast)) %>% 
  cols_label(
    feat_funding_short = "Funding source", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    diff_nice = "Posterior ∆",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_style(
    style = cell_fill(color = "grey90"),
    locations = cells_row_groups()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, diff_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior difference in marginal means is strictly positive or negative—it is the proportion of the posterior difference in marginal means that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```


## H~4c~: Relationship with host government, funding sources, and issue area

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 4~c~: Relationship with host government, funding sources, and issue area

**Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors and work in less contentious areas (emergency response and refugee relief)**
:::

## Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Few wealthy donors or Government} \mid \text{(Relationship = Under crackdown or Criticized) and (Issue = less contentious)} \right) \bigr] - \\
&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Funding} = \text{Many small donors} \mid \text{(Relationship = Under crackdown or Criticized) and (Issue = less contentious)} \right) \bigr]
\end{aligned}
$$

```{r h4c-calc-estimands}
h4c_mms <- preds_all %>% 
  group_by(feat_funding_short, feat_govt, feat_issue, .draw) %>% 
  summarize(avg = mean(.epred))

h4c_amces <- h4c_mms %>% 
  group_by(feat_govt, feat_issue) %>% 
  compare_levels(variable = avg, by = feat_funding_short, comparison = "control")
```

```{r plot-h4c-mm-amce, fig.width=10, fig.height=6.5, out.width="100%"}
#| column: body-outset-right
p_h4c_mms <- h4c_mms %>%
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_issue),
    ncol = 4,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Libre Franklin",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )
p_h4c_mms

baselines <- expand_grid(
  avg = 0,
  feat_funding_short = "Many small donors",
  feat_govt = levels(h4c_mms$feat_govt),
  feat_issue = levels(h4c_mms$feat_issue)
)

p_h4c_amces <- h4c_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_funding_short,
    delim = " - ", 
    names = c("feat_funding_short", "reference_level")
  ) %>% 
  bind_rows(baselines) %>% 
  mutate(feat_funding_short = factor(feat_funding_short, levels = levels(h4c_mms$feat_funding_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h4c_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_funding_short, fill = feat_funding_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(3, 6, 5)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_issue),
    ncol = 4,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Libre Franklin",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h4c_amces
```
