---
title: "H~5~: Organizational practices"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(scales)
library(ggh4x)
library(glue)
library(gt)
library(gtExtras)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

preds_all <- tar_read(preds_conditional_treatment_only)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

# H~5a~: Financial transparency

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~a~: Financial transparency

**Donors will show increased willingness to donate to NGOs that are financially transparent**

*Mechanism: Perception of efficacy*
:::

## Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Transparency} = \text{Yes} \right) \bigr]
\end{aligned}
$$
```{r h5a-calc-estimands}
h5a_mms <- preds_all %>% 
  group_by(feat_transp_short, .draw) %>% 
  summarize(avg = mean(.epred))

h5a_amces <- h5a_mms %>% 
  group_by(feat_transp_short) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")
```

```{r plot-h5a-mm-amce, fig.width=10, fig.height=4, out.width="100%"}
#| column: body-outset-right
p_h5a_mms <- h5a_mms %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  )

p_h5a_amces <- h5a_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_transp_short = unique(.$reference_level)) %>%
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(h5a_mms$feat_transp_short))) %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Average marginal component effects (AMCEs)"
  )

p_h5a_mms | p_h5a_amces
```

```{r table-h5a-mm-amce}
#| column: body-outset-right
h5a_tbl_mm <- h5a_mms %>% 
  group_by(feat_transp_short) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_transp_short)) %>% 
  select(feat_transp_short, nice)

h5a_tbl_amces <- h5a_amces %>% 
  group_by(feat_transp_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_transp_short = str_replace(feat_transp_short, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_transp_short)) %>% 
  select(contrast = feat_transp_short, amce_nice = nice, p_neq_0)

bind_cols(
  h5a_tbl_mm, 
  add_row(h5a_tbl_amces, contrast = "*(Reference)*")
) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, amce_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_transp_short, contrast)) %>% 
  cols_label(
    feat_transp_short = "Transparency", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    amce_nice = "Posterior AMCE",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, amce_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```


# H~5b~: Relationship with host government, organizational practices

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~b~: Relationship with host government, organizational practices

**Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent**
:::

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Transparency} = \text{Yes} \mid \text{Relationship = Under crackdown or Criticized} \right) \bigr] - \\
&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Transparency} = \text{Yes} \mid \text{Relationship = Friendly} \right) \bigr]
\end{aligned}
$$

```{r h5b-calc-estimands}
h5b_mms <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

h5b_amces <- h5b_mms %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")
```

```{r plot-h5b-mm-amce, fig.width=10, fig.height=6.5, out.width="100%"}
#| column: body-outset-right
p_h5b_mms <- h5b_mms %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_wrap(vars(feat_govt), ncol = 1) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )

p_h5b_amces <- h5b_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_transp_short = "No", feat_govt = "Friendly relationship with government") %>%
  add_row(avg = 0, feat_transp_short = "No", feat_govt = "Criticized by government") %>%
  add_row(avg = 0, feat_transp_short = "No", feat_govt = "Under government crackdown") %>%
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(h5b_mms$feat_transp_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h5b_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_wrap(vars(feat_govt), ncol = 1) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h5b_mms | p_h5b_amces
```

```{r table-h5b-mm-amce}
#| column: body-outset-right
h5b_tbl_mm <- h5b_mms %>% 
  group_by(feat_govt, feat_transp_short) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_govt), desc(feat_transp_short)) %>% 
  select(feat_transp_short, feat_govt, nice)

h5b_tbl_amces <- h5b_amces %>% 
  group_by(feat_govt, feat_transp_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  ungroup() %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_transp_short = str_replace(feat_transp_short, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_transp_short, diff_nice = nice, p_neq_0)

bind_cols(
  h5b_tbl_mm, 
  h5b_tbl_amces %>% 
    add_row(contrast = NA, .after = 1) %>% 
    add_row(contrast = NA, .after = 3) %>% 
    add_row(contrast = NA, .after = 5)
) %>% 
  mutate(feat_govt = fct_relabel(feat_govt, ~paste("Relationship with government:", .x))) %>% 
  group_by(feat_govt) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, diff_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_transp_short, contrast)) %>% 
  cols_label(
    feat_transp_short = "Transparency", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    diff_nice = "Posterior ∆",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_style(
    style = cell_fill(color = "grey90"),
    locations = cells_row_groups()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, diff_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior difference in marginal means is strictly positive or negative—it is the proportion of the posterior difference in marginal means that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```


## H~5c~: Relationship with host government, organizational practices, and funding sources

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~c~: Relationship with host government, organizational practices, and funding sources

**Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and are funded primarily by numerous small private donors**
:::

## Estimand

TODO

```{r h5c-calc-estimands}
h5c_mms <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, feat_funding_short, .draw) %>% 
  summarize(avg = mean(.epred))

h5c_amces <- h5c_mms %>% 
  group_by(feat_govt, feat_funding_short) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")
```

```{r plot-h5c-mm-amce, fig.width=10, fig.height=6.5, out.width="100%"}
#| column: body-outset-right
p_h5c_mms <- h5c_mms %>%
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_funding_short),
    ncol = 3,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )
p_h5c_mms

baselines <- expand_grid(
  avg = 0,
  feat_transp_short = "No",
  feat_govt = levels(h5c_mms$feat_govt),
  feat_funding_short = levels(h5c_mms$feat_funding_short)
)

p_h5c_amces <- h5c_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  bind_rows(baselines) %>% 
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(h5c_mms$feat_transp_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h5c_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_funding_short),
    ncol = 3,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h5c_amces
```


# H~5d~: Relationship with host government, organizational practices, and issue area

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~d~: Relationship with host government, organizational practices, and issue area

**Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief)**
:::

## Estimand

TODO

```{r h5d-calc-estimands}
h5d_mms <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, feat_issue, .draw) %>% 
  summarize(avg = mean(.epred))

h5d_amces <- h5d_mms %>% 
  group_by(feat_govt, feat_issue) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")
```

```{r plot-h5d-mm-amce, fig.width=10, fig.height=6.5, out.width="100%"}
#| column: body-outset-right
p_h5d_mms <- h5d_mms %>%
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_issue),
    ncol = 4,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )
p_h5d_mms

baselines <- expand_grid(
  avg = 0,
  feat_transp_short = "No",
  feat_govt = levels(h5d_mms$feat_govt),
  feat_issue = levels(h5d_mms$feat_issue)
)

p_h5d_amces <- h5d_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  bind_rows(baselines) %>% 
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(h5d_mms$feat_transp_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h5d_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_issue),
    ncol = 4,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h5d_amces
```


# H~5e~: Relationship with host government, organizational practices, issue area, and funding sources

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~e~: Relationship with host government, organizational practices, issue area, and funding sources

**Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) and are funded by numerous small donors**
:::


TODO

```{r h5e-calc-estimands}
h5e_mms <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, feat_funding_short, feat_issue, .draw) %>% 
  summarize(avg = mean(.epred))

h5e_amces <- h5e_mms %>% 
  group_by(feat_govt, feat_funding_short, feat_issue) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")
```

```{r plot-h5e-mm-amce, fig.width=14, fig.height=6.5, out.width="100%"}
#| column: page
p_h5e_mms <- h5e_mms %>%
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_funding_short, feat_issue),
    ncol = 12,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.margin = margin(l = -7, t = 0)
  )
p_h5e_mms

baselines <- expand_grid(
  avg = 0,
  feat_transp_short = "No",
  feat_govt = levels(h5e_mms$feat_govt),
  feat_funding_short = levels(h5e_mms$feat_funding_short),
  feat_issue = levels(h5e_mms$feat_issue)
)

p_h5e_amces <- h5e_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  bind_rows(baselines) %>% 
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(h5e_mms$feat_transp_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(h5e_mms$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_transp_short, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(10, 3)], guide = "none") +
  facet_nested_wrap(
    vars(feat_govt, feat_funding_short, feat_issue),
    ncol = 12,
    strip = strip_nested(
      text_x = list(element_text(
        family = "Clear Sans",
        face = "bold"
      ), NULL),
      background_x = list(element_rect(fill = "grey96"), NULL),
      by_layer_x = TRUE
    )
  ) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Difference in estimated marginal means"
  )

p_h5e_amces
```


# H~5f~: Accountability

::: {.callout-tip icon=false}
## {{< iconify fa6-solid flask-vial >}} Hypothesis 5~f~: Accountability

**Donors should be no more or less likely to donate to NGOs that are accountable and hold regular third party audits**

*Mechanism: Donors don't necessarily seek assurance through third-party programs/audits and charity watchdogs, but rather through word of mouth, personal scrutiny and local networks*
:::

## Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Accountability} = \text{Yes} \right) \bigr]
\end{aligned}
$$

```{r h5f-calc-estimands}
h5f_mms <- preds_all %>% 
  group_by(feat_acc_short, .draw) %>% 
  summarize(avg = mean(.epred))

h5f_amces <- h5f_mms %>% 
  group_by(feat_acc_short) %>% 
  compare_levels(variable = avg, by = feat_acc_short, comparison = "control")
```

```{r plot-h5f-mm-amce, fig.width=10, fig.height=4, out.width="100%"}
#| column: body-outset-right
p_h5f_mms <- h5f_mms %>% 
  ggplot(aes(x = avg, y = feat_acc_short, fill = feat_acc_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(9, 4)], guide = "none") +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  )

p_h5f_amces <- h5f_amces %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_acc_short,
    delim = " - ", 
    names = c("feat_acc_short", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_acc_short = unique(.$reference_level)) %>%
  mutate(feat_acc_short = factor(feat_acc_short, levels = levels(h5f_mms$feat_acc_short))) %>% 
  ggplot(aes(x = avg, y = feat_acc_short, fill = feat_acc_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(11)) +
  scale_fill_manual(values = clrs$prism[c(9, 4)], guide = "none") +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Average marginal component effects (AMCEs)"
  )

p_h5f_mms | p_h5f_amces
```

```{r table-h5f-mm-amce}
#| column: body-outset-right
h5f_tbl_mm <- h5f_mms %>% 
  group_by(feat_acc_short) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_acc_short)) %>% 
  select(feat_acc_short, nice)

h5f_tbl_amces <- h5f_amces %>% 
  group_by(feat_acc_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_acc_short = str_replace(feat_acc_short, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_acc_short)) %>% 
  select(contrast = feat_acc_short, amce_nice = nice, p_neq_0)

bind_cols(
  h5f_tbl_mm, 
  add_row(h5f_tbl_amces, contrast = "*(Reference)*")
) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, amce_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_acc_short, contrast)) %>% 
  cols_label(
    feat_acc_short = "Accountability", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    amce_nice = "Posterior AMCE",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, amce_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```
