---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Lewis & Clark College  
[Marc Dotson](https://occasionaldivergences.com/about) • Utah State University  
[Andrew Heiss](https://www.andrewheiss.com) • Georgia State University

<!-- badges: start -->
<!-- badges: end -->

---

## Abstract

Non-governmental organizations (NGOs) are a core component of global governance. Over the last two decades, however, many host governments have cracked down on NGOs using a series of laws that adversely affect the availability and use of traditional NGO funding. In response to this dramatic shift in the funding landscape, international NGOs have increasingly turned to individual donors to offset the loss funding. Prior research on NGO fundraising has examined how financial transparency and organizational accountability influence individual donor behavior. How do these elements of private donor decision-making interact with structural-level factors such as worsening host country civic environments when deciding to donate? Using a conjoint experiment with likely donors in the US, we find that transparency and accountability can protect against the dampening effect of host government crackdown and criticism of NGOs. Our results have important implications for how NGOs can adapt to worsening civic environments.


### `r emoji_glue(":zipper_mouth_face::sun_behind_large_cloud::person_walking:")`: Note on "silent skywalk" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate a [NICKA](https://en.wikipedia.org/wiki/List_of_U.S._Department_of_Defense_and_partner_code_names)-style internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.3.0 (2023-04-21).

codename(type = "nicka", seed = "why donors donate")
#> [1] "silent skywalk"
```


## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
